---
title: "Parallel Processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parallel Processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In summary, all you have to do is set `nthread = x` where `x` is the number of threads to use, usually the number of CPU cores to let the engine use from you own computer. E.g. `set_engine("lightgbm", nthread = 8)`.

Parsnip principles states that *parallel processing must be explictly requested by the user*^[https://tidymodels.github.io/model-implementation-principles/parallel-processing.html], so if `nthread` were not specified, just a single thread will be used.

```{r, eval = FALSE, message=FALSE}
library(tidymodels)
library(treesnip)
big_mtcars <- mtcars %>% sample_n(1000000, replace = TRUE)

# model spec
model <- boost_tree(tree_depth = tune()) %>%
  set_mode("regression") %>%
  set_engine("catboost", nthread = 8)
  # set_engine("lightgbm", nthread = 8)
  # set_engine("xgboost", nthread = 8)

# recipe
rec <- recipe(mpg ~ ., data = big_mtcars)

# workflows
wf <- workflow() %>% add_recipe(rec) %>% add_model(model)
tune_grid(wf, vfold_cv(big_mtcars, v = 2))
```

PS: originally, there is no `nthread` parameter for {catboost} but {treesnip} made `nthread` an alias for `thread_count` for convenience.

