---
title: "Parallel Processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parallel Processing}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Rules of thumb

- When tuning: `doParallel::registerDoParallel(8)` then `tune_grid()`/`tune_bayes()`
- When fitting: `set_engine("catboost", nthread = 8)` then `fit()`/`last_fit()`


## Packages and data

```{r data, eval = TRUE, message=FALSE}
library(tidymodels)
library(treesnip)

big_mtcars <- mtcars %>% sample_n(100000, replace = TRUE)
```

## Forking (workers)

Set `doParallel::registerDoParallel(8)` before `tune_grid()` or `tune_bayes()` and you are good to go. Just a reminder to watch out the memory usage. Forking can make your machine run out of memory.

```{r forking, cache=TRUE, eval = FALSE}
# model spec
model <- boost_tree(tree_depth = tune()) %>%
  set_mode("regression") %>%
  set_engine("catboost")
# set_engine("lightgbm")

# recipe
rec <- recipe(mpg ~ ., data = big_mtcars)

# workflows
wf <- workflow() %>% add_recipe(rec) %>% add_model(model)

# parallel tune grid
doParallel::registerDoParallel(8)
tg <- tune_grid(wf, vfold_cv(big_mtcars, 5))
```

## Threading

In summary, all you have to do is set `nthread = x` where `x` is the number of threads to use, usually the number of CPU cores to let the engine use from you own computer. E.g. `set_engine("lightgbm", nthread = 8)`.

Parsnip principles states that *parallel processing must be explictly requested by the user*^[https://tidymodels.github.io/model-implementation-principles/parallel-processing.html], so if `nthread` were not specified, just a single thread will be used.
`
```{r threading, eval = FALSE, message=FALSE}
model <- boost_tree(tree_depth = tune()) %>%
  set_mode("regression") %>%
  set_engine("catboost", nthread = 8) %>%
  finalize_model(select_best(tg, "rmse"))

wf <- workflow() %>% add_recipe(rec) %>% add_model(model)

definitive_fit <- fit(wf,  big_mtcars)
```

PS: originally, there is no `nthread` parameter for {catboost} but {treesnip} made `nthread` an alias for `thread_count` for convenience.



## Tune Grid Benchmark (speed)

### Findings:

- Threading is better for small datasets
- Forking is better for large datasets
- Memory is an issue when forking

### Parameters (all combinations of...):

- **the size of the data.frame:** 1.000 vs 1.000.000 rows
- **the number of models trained in grid search:** 48 vs 1875
- **threads/workers:** 1/8, 2/4, 4/2 and 8/1 (there was a constraint of `workers*threads = 8`)

### Machine: 

- Intel Core i7-8565U 
- CPU @ 1.80GHz × 8 cores 
- Dell Inspiron
- RAM 16GB
- Linux Ubuntu 20.04 64bit


**further reading:** [Parallel Processing from Applied Predictive Modeling (Kuhn and Johnson, 2014)](http://appliedpredictivemodeling.com/blog/2018/1/17/parallel-processing) 

### Code:

```{r code, cache=TRUE, eval = FALSE}
library(tidymodels)
library(treesnip)

# main function to tune_grid with given benchmark parameters
execute_tune_grid <- function(df, threads, workers, cv_folds, tune_grid, engine = "lightgbm") {
  ##############################################################
  df_splits <- vfold_cv(df, v = cv_folds)
  df_rec <- recipe(mpg ~ ., data = df) 
  df_model <- boost_tree(
    mtry = 3, 
    trees = 600, 
    sample_size = 0.7,
    min_n = tune(),
    tree_depth = tune(),
    learn_rate = tune(),
    loss_reduction = tune()
  ) %>% 
    set_mode("regression") %>% 
    set_engine(as.character(engine), nthread = threads)
  
  df_wf <- workflow() %>% 
    add_model(df_model) %>% 
    add_recipe(df_rec)
  
  ##############################################################
  doParallel::registerDoParallel(workers)
  tg <- tune_grid(df_wf, df_splits, grid = tune_grid)
  
  tg
}

# parameters
parameters <- expand.grid(
  engine = "lightgbm",
  cv_folds = 8,
  tune_grid = c(2, 5, 10),
  nrow = c(1e2, 1e4, 1e6),
  threads = c(1, 2, 4, 8)
) %>%
  mutate(
    workers = 8 %/% threads
  ) %>%
  bind_rows(
    expand.grid(
      engine = "lightgbm",
      cv_folds = 8,
      tune_grid = c(2, 5, 10),
      nrow = c(1e2, 1e4, 1e6),
      threads = 1,
      workers = 1
    )
  )

# bench::mark
set.seed(1)
bm <- bench::press(
  {
    Sys.sleep(2)
    df <- mtcars %>% sample_n(nrow, replace = TRUE)
    bench::mark(    
      execute_tune_grid(df, threads, workers, cv_folds, tune_grid, engine),
      check = FALSE,
      iterations = 3,
      memory = FALSE,
      filter_gc = FALSE,
      time_unit = "m"
    )
  },
  .grid = parameters
)
```

### Results:

```{r, include = FALSE}
bm <- readr::read_rds(system.file("benchmark_parallel_processing_vignette.rds", package = "treesnip"))
```


```{r result, message=FALSE, warning=FALSE, fig.width=8}
bm %>%
  mutate(
    nrow = paste("nrows:", nrow),
    models_trained = (tune_grid ^ 4) * cv_folds,
    max = map_dbl(time, max)/60,
    threads = glue::glue("Threads: {threads} (workers: {workers})")
  ) %>%
  ggplot(aes(x = models_trained, y = median, colour = factor(threads))) +
  facet_wrap(~nrow, scales = "free_y") +
  geom_pointrange(aes(ymin = min, ymax = max)) +
  geom_line() +
  labs(x = "Models trained", y = "Time elapsed (minutes)", colour = "Threads (Workers)")
```


