---
title: "Threading vs Forking Benchmark - Tune Grid Speed"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{threading-forking-benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Findings:

- Forking better for small datasets
- Combination of Forking and Threading is better for large datasets


```{r, include = FALSE}
library(tidymodels)
library(treesnip)
# saveRDS(bm, "inst/benchmark_parallel_processing_vignette.rds")
bm <- readr::read_rds(system.file("benchmark_parallel_processing_vignette.rds", package = "treesnip"))
```


```{r result, message=FALSE, warning=FALSE, fig.width=8, echo = FALSE, fig.height=3}
bm %>%
  mutate(
    nrow = paste("nrows:", scales::number(nrow, 1)),
    models_trained = (tune_grid ^ 4) * cv_folds,
    max = map_dbl(time, max)/60,
    threads = glue::glue("Threads: {threads} (workers: {workers})")
  ) %>%
  ggplot(aes(x = models_trained, y = median, colour = factor(threads))) +
  facet_wrap(~nrow, scales = "free_y") +
  geom_pointrange(aes(ymin = min, ymax = max)) +
  geom_line() +
  labs(x = "Models trained", y = "Time elapsed (minutes)", colour = "Threads (Workers)", title = "engine:lightgbm - CV-folds:3 - hyperparameters: 4- trees: 600")
```

## Parameters (all combinations of...):

- **the size of the data.frame:** 1.000 vs 1.000.000 rows
- **the number of models trained in grid search:** 48 vs 1875
- **threads/workers:** 1/8, 2/4, 4/2 and 8/1 (there was a constraint of `workers*threads = 8`)

## Machine: 

- Intel Core i7-8565U 
- CPU @ 1.80GHz × 8 cores 
- Dell Inspiron
- RAM 16GB
- Linux Ubuntu 20.04 64bit

## Code:
```{r, message=FALSE}
library(tidymodels)
library(treesnip)
```

```{r code, cache=TRUE, eval = FALSE}
# main function to tune_grid with given benchmark parameters
execute_tune_grid <- function(df, threads, workers, cv_folds, tune_grid, engine = "lightgbm") {
  ##############################################################
  df_splits <- vfold_cv(df, v = cv_folds)
  df_rec <- recipe(mpg ~ ., data = df) 
  df_model <- boost_tree(
    mtry = 3, 
    trees = 600, 
    sample_size = 0.7,
    min_n = tune(),
    tree_depth = tune(),
    learn_rate = tune(),
    loss_reduction = tune()
  ) %>% 
    set_mode("regression") %>% 
    set_engine(as.character(engine), nthread = threads)
  
  df_wf <- workflow() %>% 
    add_model(df_model) %>% 
    add_recipe(df_rec)
  
  ##############################################################
  doParallel::registerDoParallel(workers)
  tg <- tune_grid(df_wf, df_splits, grid = tune_grid)
  
  tg
}

# parameters
parameters <- expand.grid(
  engine = "lightgbm",
  cv_folds = 3,
  tune_grid = c(2, 5),
  nrow = c(1e3, 1e6),
  threads = c(1, 2, 4, 8)
) %>%
  mutate(
    workers = 8 %/% threads
  )

# bench::mark
set.seed(1)
bm <- bench::press(
  {
    Sys.sleep(3)
    df <- mtcars %>% sample_n(nrow, replace = TRUE)
    bench::mark(    
      execute_tune_grid(df, threads, workers, cv_folds, tune_grid, engine),
      check = FALSE,
      iterations = 3,
      memory = FALSE,
      filter_gc = FALSE,
      time_unit = "m"
    )
  },
  .grid = parameters
)
```

